### 2022-04-14 abstract.tex
copyedits, add desc stat table

My final article takes a deeper dive into the formation of social norms
governing health behaviors in cases of extreme uncertainty. I specifically use
the cases of stay-at-home rates and vaccination rates as responses to public
health recommendations to mitigate the COVID-19 pandemic. Using the theories of
associative diffusion and the integrated theoretical framework of norms, I test
models of behavioral adaption to public health recommendations and patterns of
complex contagion (the need for repeated exposures to a something novel for it
to diffuse) using linear mixed effects models. My results show that complex
contagion is a valid framework for the social contagion of new norms during
COVID-19. However, I find an important novel moderating effect of signal
discordance; if there is diversity in the information received by an ego,
contagion is less likely to occur. This paper shows that the contagion process
is not fully understood without looking at the context of each exposure to a
contagion within the range of contagions one experiences.

>>>>

My final article takes a deeper dive into the formation of social norms
governing health behaviors in cases of extreme uncertainty. I specifically use
the cases of stay-at-home rates and vaccination rates as responses to public
health recommendations to mitigate the COVID-19 pandemic. Using the theories of
associative diffusion and the integrated theoretical framework of norms, I test
models of behavioral adaption to public health recommendations and patterns of
complex contagion (the need for repeated exposures to something novel for it
to diffuse) using linear mixed effects models. My results show that complex
contagion is a valid framework for the social contagion of new norms during
COVID-19. However, I find an important novel moderating effect of signal
discordance; if there is diversity in the information received by an ego,
contagion is less likely to occur. This paper shows that the contagion process
is not fully understood without looking at the context of each exposure to a
contagion within the range of contagions one experiences.

### 2022-04-14 ch1.tex
copyedits, add desc stat table

A critical turning point in the study of the flow of information was put forth
by renowned sociologist Paul Lazarsfeld and colleagues Verelson and Gaudet
\citeyearpar{lazarsfeldPeopleChoice1944}. They theorized that the media did not
directly influence the masses, but that a "Two-Step Flow" was occurring where
"opinion leaders" interpreted messages from mass media and then, in turn,
promoted their opinions to the masses \citep{katzPersonalInfluencePart1955}. The
opinion leaders interpret, explain, and diffuse contents to "opinion followers".
From a networks perspective, "people rarely act on mass-media information unless
it is also transmitted through personal ties" \citep[p.
1374]{granovetterStrengthWeakTies1973}. Lazarsfeld et al's
\citeyearpar{lazarsfeldPeopleChoice1944} first study focused on politics, but
follow up studies by \citet{katzPersonalInfluencePart1955} showed that this
model matched information flows beyond the political realm. In contrast to the
one-step flow, where media directly influences the masses, Lazarsfeld's theory
remains in prominence today. While many scholars assumed that modern
micro-targeted marketing strategies would provide evidence for a one-step flow
perspective \citep{bennettOneStepFlowCommunication2006}, many scholars find that
"opinion leaders" (often influencers and celebrities) still create a two-step
flow of information transmission \citep{choi15, hilbertOneStepTwo2017}.

The spread of information has been thoroughly studied from the perspective of
social networks. Individuals engage with each other and their distributed ties
to create community contexts where norms, beliefs, and values circulate.
Information diffuses through communities and social networks
\citep{fowler2010cooperative, bond_etal12, klarEffectNetworkStructure2017}.
However, "different things spread in different ways and to different extents"
\citep[p. 563]{christakisSocialContagionTheory2013} and information can spread
differently than a virus or behavior. Important studies in the spread of
information through networks include the study of adoption of innovations by
physicians \citep{colemanDiffusionInnovationPhysicians1957}, how new inventions
are adopted from early adopters to laggards
\citep{rogersDiffusionInnovations1962}, how weak ties help bind networks and
diffuse information \citep{granovetterStrengthWeakTies1973}, how social ties
spread information about job opportunities
\citep{granovetterGettingJobStudy1995, montgomeryJobSearchNetwork1992}, how
information spreads best when weak and strong ties are balanced (embeddedness)
\citep{uzziSocialStructureCompetition1997}, and more recent innovations in the
stickiness of complex contagion \citep{centolaComplexContagionsWeakness2007}.

@@

The introduction of the internet and digital technologies introduced an epochal
disruption in the communication and informational-search mediums that people had
available to them. Computer-mediated communication was the largest communication
revolution in a century, changing the speed and availability of communication
between individuals by electronic mail, instant messaging, two-way interactive
video calls, discussion forums, blogs, and social networking
\citep{rainie2012networked}. Computer-mediated communication encompasses any
sort of communication through computers connected to the internet; the umbrella
term involves both synchronous and asynchronous communication as well as
one-to-one, one-to-many, or many-to-many exchanges. Social media, on the other
hand, are more specific, and are "Internet-based, disentrained, and persistent
channels of mass personal communication facilitating perceptions of interactions
among users, deriving value primarily from user-generated content” \citep[][p.
50]{carr2015social}. While an email between colleagues is computer-mediated
communication, it is not social media as there is only one interaction between
users and there is no value derived from the user-generated content outside the
pair \citep{bayer_etal20}. Even more restrictive still is the social networking
site, which consists of a profile, a network, and a stream or a feed
\citep{boyd2007social, ellison2013sociality},  often also involving a messaging
component  \citep{bayer_etal20}. Social networking sites are also interesting
because they often uniquely combine one-to-one communication within the context
of one-to-many communication, like resharing a television news story to your
Facebook feed and chatting with others about the content in the comments
section. But what do these classifications mean for the information-seeking
behaviors of individuals in the 21st century? Not only is it quicker to perform
network activation and seek support, but people are "permanently online,
permanently connected" making activation easy \citep{vorderer2017permanently}.
The constant connectedness facilitates communication, but also presents people
with overchoice, or the difficulty in deciding because simply every network
connection you may want to activate is permanently connected and easy to reach.
In other words, constant connection may be debilitating to many who are seeking
information. \citet{smallSomeoneTalk2017} provides many examples of this when he
demonstrates that often, graduate students who needed support simply asked
"miscellaneous classmates encountered down the hall" (p. 176) rather than their
trusted, long-term confidants. It seems that network activation may rely on more
than just trust; namely: convenience.

@@

One way to find information is to activate network ties to find out information
through a form of social support. Social support, while previously used
interchangeably with the term social networks and social integration
\citep{houseStructuresProcessesSocial1988}, consisted of the emotional,
informational, and instrumental assistance functions performed by social ties.
Social support has strong and measurable association with health outcomes
\citep{houseMeasuresConceptsSocial1985, thoitsMechanismsLinkingSocial2011}.
Informational support is the process of seeking "help in defining,
understanding, and coping with problematic events and include education, advice,
or referral to another source of support" \citep[p. 640]{winemiller_etal93}.
Brashers, a health communications researcher, defines informational support
slightly differently, focusing on the exchange of information that "facilitates
coping with life stresses... that may be exchanged among members of a support
network" \citeyearpar[p. 260]{brashersInformationSeekingAvoiding2002}. Both
definitions provide important lenses for my question of how computer-mediated or
interpersonal information-seeking strategies vary. Some research has revealed
that Facebook users who utilize the platform to find information and to maintain
relationships have high levels of bridging social capital \citep{liu2016meta}.
However, this leaves the question of who activates their social networks for
information using social networking sites like Facebook instead of activating
these networks in face-to-face interactions.

@@

There has been some research into how different platforms allow for different
user experiences or "affordances" \citep{boyd2010social}. For example,
\citet{bossetta18} first show that all social media platforms must provide
"tools that are easy to use and functional to the demands of varying user
demographics" (p. 473). They also explain how each digital architecture enables,
constrains and shapes how users behave on their platform and in virtual space,
leading to many different experiences along the dimensions of searchability,
connectivity, and privacy.

@@

Social norms form the building blocks of social organization and have been a
focus of sociology since the beginning. In a 2020 Annual Review of Sociology
piece focused on social norms, Horne and Mollborn define norms as "group-level
evaluations of behavior ... [or] when people have expectations about how others
evaluate behaviors" \citeyearpar[p. 468-69]{horneNormsIntegratedFramework2020}.
One of the principal architects of sociology, \'{E}mile Durkheim, saw that
society exerted powerful forces on individuals and classified people's norms,
beliefs, and values as parts of collective consciousness that provide for
societal integration \citeyearpar{durkheimSuicide1897,
durkheimDivisionLaborSociety1933}. For Durkheim, norms were therefore the glue
that held society together in cohesion.

@@

Norms are formed inside of social networks; "Networks operate in a larger
cultural context that can facilitate or inhibit acceptance of general cultural
norms and beliefs" \citep[][p. 44]{pescosolidoDurkheimSuicideReligion1989}.
Because of this, people tend to resemble their network as a form of network
autocorrelation and homophily \citep{dellapostaWhyLiberalsDrink2015}. While
"birds of a feather flock together" in homophily
\citep{mcphersonBirdsFeatherHomophily2001}, attitudes and behaviors are shaped
by peers, creating filter bubbles of social influence
\citep{dellapostaWhyLiberalsDrink2015}.

\citet{goldbergSocialContagionAssociative2018} argue that the state of the
literature on diffusion and social contagion, discussed above, is erroneous.
They propose an important linkage that had been missed in previous research of
the diffusion of behavior (or, as I interpret it, the creation of new norms).
They argue that what actually diffuses during social contagion are the
perceptions about which beliefs or behaviors are compatible with one another,
what they call "associative diffusion". This relates back to the theory of
relational norms, where individuals assume what their peers will approve and
disapprove of. For these assumptions to take place, they must understand which
behaviors and beliefs are possessed by whom in their network and how new
associations may influence or disrupt the expected actions. In times of
behavioral uncertainty, I theorize that individuals strategically look towards
other high-status individuals and institutions, and perform regimented searches
to establish for themselves the behavior that is compatible with other cognitive
associations and biases they hold.

@@

Computational social science and social science disciplines are quick to adopt
new sources of digital trace data for use in academic research. In this article,
I examine the criterion validity of one of these new sources, Google Search
Trends. I validate how Google Trends can be used to demonstrate three different
cases, namely attitudes, disease prevalence and political preferences using five
different validated data sources. This paper asks, "How can we operationalize
Google Search Trends as a valid indicator for uses in social science research?"
I use Pearson and Repeated Measures Correlations between the Google Trends and
the validated indicators as well as multiple linear regression (for
cross-sectional datasets) and fixed effect hierarchical linear models (for
longitudinal datasets) as additional tests of the data.

@@

My final article takes a deeper dive into the formation of social norms
governing health behaviors in cases of extreme uncertainty using the cases of
both stay-at-home rates and vaccination rates as responses to public health
recommendations to mitigate the COVID-19 pandemic. Using theories like
associative diffusion and the integrated theoretical framework of norms, I test
models of behavioral adaption to public health recommendations and patterns of
complex contagion (the need for repeated exposures to a something novel for it
to diffuse) using linear mixed effects models.

These models show that complex contagion is a valid framework for the social
contagion of new norms during COVID-19. Importantly, I find a novel moderating
effect of signal discordance, the contextual diversity of signals received by an
ego. If there is diversity in the information received by an ego, contagion is
less likely to occur. This paper shows that the contagion process cannot be not
fully understood without looking at the context of each exposure to a contagion
within the range of contagions one experiences.

@@

% micro behaviors to macro processes
Bridging the contribution of search as an agentic process and critical big data
studies, this dissertation contributes to the question posed by
\citet{breigerScaling2015}, "what are the behavioral processes that lead to
macro-level outcomes?" With the exponentially growing size of digital-trace data
that is available, it is important to not only uncover the external and
criterion validity of the data and correspondence to real world behaviors (see
\hyperlink{paper-1}{Article 1}) but also important to understand the social
forces that lead to the big data. One of the contributions of
\hyperlink{paper-2}{Article 2} is to dive deep into this question, uncovering
why some people choose to use web search engines over other information search
modalities. These papers are a first step in validating and establishing a
deeper understanding of who the users are that contribute to macro-level
aggregate trends.

% social norms
Additionally, this dissertation investigates how social norms are formed in
situations of uncertainty, an area that is far less developed in sociology.
While sociologists have studied social norms since the formation of the
discipline, norms are most often studied under the lens of the emergence of
norms and transmission of norms between groups, deviance and social control, the
interaction of norms as "mental representations of appropriate behavior"
\citep{aarts2003silence}, and research categorizing different types of norms
(e.g., descriptive versus injunctive, prescriptive versus proscriptive). The
focus on the formation of new social norms during times of uncertainty has
implications for interventions in social behavioral outcomes, research on
cultural change, and the spread of cognitive associations. Moreover,
\hyperlink{paper-3}{Article 3} focuses on how conflicting behaviors become
settled into cognitive associations, providing a relevant framework for research
on unrest and disasters.

>>>>

A critical turning point in the study of the flow of information was put forth
by renowned sociologist Paul Lazarsfeld and colleagues Verelson and Gaudet
\citeyearpar{lazarsfeldPeopleChoice1944}. They theorized that the media did not
directly influence the masses, but that a ``Two-Step Flow'' was occurring where
``opinion leaders'' interpreted messages from mass media and then, in turn,
promoted their opinions to the masses \citep{katzPersonalInfluencePart1955}. The
opinion leaders interpret, explain, and diffuse contents to ``opinion followers''.
From a networks perspective, ``people rarely act on mass-media information unless
it is also transmitted through personal ties'' \citep[p.
1374]{granovetterStrengthWeakTies1973}. Lazarsfeld et al's
\citeyearpar{lazarsfeldPeopleChoice1944} first study focused on politics, but
follow up studies by \citet{katzPersonalInfluencePart1955} showed that this
model matched information flows beyond the political realm. In contrast to the
one-step flow, where media directly influences the masses, Lazarsfeld's theory
remains in prominence today. While many scholars assumed that modern
micro-targeted marketing strategies would provide evidence for a one-step flow
perspective \citep{bennettOneStepFlowCommunication2006}, many scholars find that
``opinion leaders'' (often influencers and celebrities) still create a two-step
flow of information transmission \citep{choi15, hilbertOneStepTwo2017}.

The spread of information has been thoroughly studied from the perspective of
social networks. Individuals engage with each other and their distributed ties
to create community contexts where norms, beliefs, and values circulate.
Information diffuses through communities and social networks
\citep{fowler2010cooperative, bond_etal12, klarEffectNetworkStructure2017}.
However, ``different things spread in different ways and to different extents''
\citep[p. 563]{christakisSocialContagionTheory2013} and information can spread
differently than a virus or behavior. Important studies in the spread of
information through networks include the study of adoption of innovations by
physicians \citep{colemanDiffusionInnovationPhysicians1957}, how new inventions
are adopted from early adopters to laggards
\citep{rogersDiffusionInnovations1962}, how weak ties help bind networks and
diffuse information \citep{granovetterStrengthWeakTies1973}, how social ties
spread information about job opportunities
\citep{granovetterGettingJobStudy1995, montgomeryJobSearchNetwork1992}, how
information spreads best when weak and strong ties are balanced (embeddedness)
\citep{uzziSocialStructureCompetition1997}, and more recent innovations in the
stickiness of complex contagion \citep{centolaComplexContagionsWeakness2007}.

@@

The introduction of the internet and digital technologies introduced an epochal
disruption in the communication and informational-search mediums that people had
available to them. Computer-mediated communication was the largest communication
revolution in a century, changing the speed and availability of communication
between individuals by electronic mail, instant messaging, two-way interactive
video calls, discussion forums, blogs, and social networking
\citep{rainie2012networked}. Computer-mediated communication encompasses any
sort of communication through computers connected to the internet; the umbrella
term involves both synchronous and asynchronous communication as well as
one-to-one, one-to-many, or many-to-many exchanges. Social media, on the other
hand, are more specific, and are ``Internet-based, disentrained, and persistent
channels of mass personal communication facilitating perceptions of interactions
among users, deriving value primarily from user-generated content'' \citep[][p.
50]{carr2015social}. While an email between colleagues is computer-mediated
communication, it is not social media as there is only one interaction between
users and there is no value derived from the user-generated content outside the
pair \citep{bayer_etal20}. Even more restrictive still is the social networking
site, which consists of a profile, a network, and a stream or a feed
\citep{boyd2007social, ellison2013sociality},  often also involving a messaging
component  \citep{bayer_etal20}. Social networking sites are also interesting
because they often uniquely combine one-to-one communication within the context
of one-to-many communication, like resharing a television news story to your
Facebook feed and chatting with others about the content in the comments
section. But what do these classifications mean for the information-seeking
behaviors of individuals in the 21st century? Not only is it quicker to perform
network activation and seek support, but people are ``permanently online,
permanently connected'' making activation easy \citep{vorderer2017permanently}.
The constant connectedness facilitates communication, but also presents people
with overchoice, or the difficulty in deciding because simply every network
connection you may want to activate is permanently connected and easy to reach.
In other words, constant connection may be debilitating to many who are seeking
information. \citet{smallSomeoneTalk2017} provides many examples of this when he
demonstrates that often, graduate students who needed support simply asked
``miscellaneous classmates encountered down the hall'' (p. 176) rather than their
trusted, long-term confidants. It seems that network activation may rely on more
than just trust; namely: convenience.

@@

One way to find information is to activate network ties to find out information
through a form of social support. Social support, while previously used
interchangeably with the term social networks and social integration
\citep{houseStructuresProcessesSocial1988}, consisted of the emotional,
informational, and instrumental assistance functions performed by social ties.
Social support has strong and measurable association with health outcomes
\citep{houseMeasuresConceptsSocial1985, thoitsMechanismsLinkingSocial2011}.
Informational support is the process of seeking ``help in defining,
understanding, and coping with problematic events and include education, advice,
or referral to another source of support'' \citep[p. 640]{winemiller_etal93}.
Brashers, a health communications researcher, defines informational support
slightly differently, focusing on the exchange of information that ``facilitates
coping with life stresses... that may be exchanged among members of a support
network'' \citeyearpar[p. 260]{brashersInformationSeekingAvoiding2002}. Both
definitions provide important lenses for my question of how computer-mediated or
interpersonal information-seeking strategies vary. Some research has revealed
that Facebook users who utilize the platform to find information and to maintain
relationships have high levels of bridging social capital \citep{liu2016meta}.
However, this leaves the question of who activates their social networks for
information using social networking sites like Facebook instead of activating
these networks in face-to-face interactions.

@@

There has been some research into how different platforms allow for different
user experiences or ``affordances'' \citep{boyd2010social}. For example,
\citet{bossetta18} first show that all social media platforms must provide
``tools that are easy to use and functional to the demands of varying user
demographics'' (p. 473). They also explain how each digital architecture enables,
constrains and shapes how users behave on their platform and in virtual space,
leading to many different experiences along the dimensions of searchability,
connectivity, and privacy.

@@

Social norms form the building blocks of social organization and have been a
focus of sociology since the beginning. In a 2020 Annual Review of Sociology
piece focused on social norms, Horne and Mollborn define norms as ``group-level
evaluations of behavior ... [or] when people have expectations about how others
evaluate behaviors'' \citeyearpar[p. 468-69]{horneNormsIntegratedFramework2020}.
One of the principal architects of sociology, \'{E}mile Durkheim, saw that
society exerted powerful forces on individuals and classified people's norms,
beliefs, and values as parts of collective consciousness that provide for
societal integration \citeyearpar{durkheimSuicide1897,
durkheimDivisionLaborSociety1933}. For Durkheim, norms were therefore the glue
that held society together in cohesion.

@@

Norms are formed inside of social networks; ``Networks operate in a larger
cultural context that can facilitate or inhibit acceptance of general cultural
norms and beliefs'' \citep[][p. 44]{pescosolidoDurkheimSuicideReligion1989}.
Because of this, people tend to resemble their network as a form of network
autocorrelation and homophily \citep{dellapostaWhyLiberalsDrink2015}. While
``birds of a feather flock together'' in homophily
\citep{mcphersonBirdsFeatherHomophily2001}, attitudes and behaviors are shaped
by peers, creating filter bubbles of social influence
\citep{dellapostaWhyLiberalsDrink2015}.

\citet{goldbergSocialContagionAssociative2018} argue that the state of the
literature on diffusion and social contagion, discussed above, is erroneous.
They propose an important linkage that had been missed in previous research of
the diffusion of behavior (or, as I interpret it, the creation of new norms).
They argue that what actually diffuses during social contagion are the
perceptions about which beliefs or behaviors are compatible with one another,
what they call ``associative diffusion''. This relates back to the theory of
relational norms, where individuals assume what their peers will approve and
disapprove of. For these assumptions to take place, they must understand which
behaviors and beliefs are possessed by whom in their network and how new
associations may influence or disrupt the expected actions. In times of
behavioral uncertainty, I theorize that individuals strategically look towards
other high-status individuals and institutions, and perform regimented searches
to establish for themselves the behavior that is compatible with other cognitive
associations and biases they hold.

@@

Computational social science and social science disciplines are quick to adopt
new sources of digital trace data for use in academic research. In this article,
I examine the criterion validity of one of these new sources, Google Search
Trends. I validate how Google Trends can be used to demonstrate three different
cases, namely attitudes, disease prevalence and political preferences using five
different validated data sources. This paper asks, ``How can we operationalize
Google Search Trends as a valid indicator for uses in social science research?''
I use Pearson and Repeated Measures Correlations between the Google Trends and
the validated indicators as well as multiple linear regression (for
cross-sectional datasets) and fixed effect hierarchical linear models (for
longitudinal datasets) as additional tests of the data.

@@

My final article takes a deeper dive into the formation of social norms
governing health behaviors in cases of extreme uncertainty using the cases of
both stay-at-home rates and vaccination rates as responses to public health
recommendations to mitigate the COVID-19 pandemic. Using theories like
associative diffusion and the integrated theoretical framework of norms, I test
models of behavioral adaption to public health recommendations and patterns of
complex contagion (the need for repeated exposures to something novel for it
to diffuse) using linear mixed effects models.

These models show that complex contagion is a valid framework for the social
contagion of new norms during COVID-19. Importantly, I find a novel moderating
effect of signal discordance, the contextual diversity of signals received by an
ego. If there is diversity in the information received by an ego, contagion is
less likely to occur. This paper shows that the contagion process cannot be
fully understood without looking at the context of each exposure to a contagion
within the range of contagions one experiences.

@@

% micro behaviors to macro processes
Bridging the contribution of search as an agentic process and critical big data
studies, this dissertation contributes to the question posed by
\citet{breigerScaling2015}, ``what are the behavioral processes that lead to
macro-level outcomes?'' With the exponentially growing size of digital-trace data
that is available, it is important to not only uncover the external and
criterion validity of the data and correspondence to real world behaviors (see
\hyperlink{paper-1}{Article 1}) but also important to understand the social
forces that lead to the big data. One of the contributions of
\hyperlink{paper-2}{Article 2} is to dive deep into this question, uncovering
why some people choose to use web search engines over other information search
modalities. These papers are a first step in validating and establishing a
deeper understanding of who the users are that contribute to macro-level
aggregate trends.

% social norms
Additionally, this dissertation investigates how social norms are formed in
situations of uncertainty, an area that is far less developed in sociology.
While sociologists have studied social norms since the formation of the
discipline, norms are most often studied under the lens of the emergence of
norms and transmission of norms between groups, deviance and social control, the
interaction of norms as ``mental representations of appropriate behavior''
\citep{aarts2003silence}, and research categorizing different types of norms
(e.g.\, descriptive versus injunctive, prescriptive versus proscriptive). The
focus on the formation of new social norms during times of uncertainty has
implications for interventions in social behavioral outcomes, research on
cultural change, and the spread of cognitive associations. Moreover,
\hyperlink{paper-3}{Article 3} focuses on how conflicting behaviors become
settled into cognitive associations, providing a relevant framework for research
on unrest and disasters.
